---
title: "P8105_hw3_as5697"
author: "Apoorva Srinivasan"
date: "10/14/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)

library(tidyverse)
library(readxl)
library(p8105.datasets)
library(ggplot2)
library(ggridges)
library(patchwork)
library(viridis)


theme_set(theme_bw() + theme(legend.position = "bottom"))
```


#Problem 1

###1.1

```{r}
data(brfss_smart2010)
```

```{r tidy_brfss}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  rename(state = locationabbr, county = locationdesc) %>%
  mutate(response = as.factor(response)) %>% 
  mutate(
    response = fct_relevel(response, c("Excellent", "Very good", "Good", "Fair", "Poor"))
    )
```


###1.2

In 2002, which states were observed at 7 locations?


```{r}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n_locations = n_distinct(county)) %>% 
  filter(n_locations == 7) %>%
  head()
```

###1.3

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.


```{r p1.3}
brfss %>%
  group_by(year, state) %>%
  summarize(num_locations = n()) %>%
  ggplot(aes(x = year, y = num_locations, color = state)) + 
  labs(
    title = "Spaghetti plot of number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "Number of locations") +
  geom_line() + 
  theme(legend.position = "right") +
  scale_color_viridis(discrete = TRUE) 
```


###1.4

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.


```{r p1.4}
brfss %>%
  spread(key = response, value = data_value) %>%
  janitor::clean_names() %>%
  filter(year == "2002" | year == "2006" | year == "2010") %>% 
  group_by(year, state) %>%
  summarize(excellent_mean = mean(excellent, na.rm = TRUE), 
            excellent_SD = sd(excellent, na.rm = TRUE)) %>%
  filter(state == "NY") %>%
  knitr::kable(digits = 1)
```


###1.5

For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.


```{r p1.5}
brfss %>% 
  group_by(year, state, response) %>% 
  summarize(average = mean(data_value, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = average, group = year, color = response)) + 
  geom_boxplot() + 
  facet_grid(~ response) +
  labs(title = "Distribution of State Level Averages Over Time", x = "Year", y = "Mean Proportions") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```


#PROBLEM 2

```{r}
data("instacart")
  instacart = instacart %>% janitor::clean_names()
```

The dataset has `r nrow(instacart)` by `r ncol(instacart)` columns. The dataset contains some key variables such as aisle , department  and product name. Each row is a single product from an order with key variables corresponding information on the `product_name`, `department`, grocery `aisle`, if it was `reordered`, day of the week and hour it was ordered, and the `user_id` who ordered it.There are `r nrow(distinct(instacart, product_id))` products that were ordered by `r nrow(distinct(instacart, user_id))` users, from `r nrow(distinct(instacart, department_id))` distinct departments.


How many aisles are there, and which aisles are the most items ordered from?

```{r}

instacart %>%
  group_by(aisle) %>%
  summarize(order_per_aisle = n()) %>%
  arrange(desc(order_per_aisle)) %>%
  head() %>%
  knitr::kable()
```

There are `r nrow(distinct(instacart, aisle))` distinct aisles, with the top 6 aisles being fresh vegetables, fresh fruits, packaged vegetables fruits, yogurt, packaged cheese, and water seltzer sparkling water.

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  mutate(aisle = reorder(aisle, desc(n_items))) %>% 
  ggplot(aes(x = aisle, y = n_items)) + geom_point() + labs(main = "Number of items ordered from each aisle", x = "Aisle", y = "No. of items")  + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(n_prod = n()) %>% 
  filter(n_prod == max(n_prod)) %>% 
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).



```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  mutate(order_dow = ordered(order_dow, levels = c(0:6), labels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean = round(mean(order_hour_of_day))) %>% 
  spread(key = order_dow, value = mean) %>% 
  knitr::kable()
```

Sunday is taken as 0. 
Both coffee ice cream and pink lady apples are mostly ordered during around noon and after.





